\sffamily An EEG-based visual Brain-Computer Interface for gaze-free communication

### 2024-11-19 rebuttal.tex

\reply{Adalberto Simeone}
Major comments:
\begin{enumerate}
	\item \todo{discuss in limitations the impact of letters and future
	work experiments with an actual usable interface}
	\todo{mention here plans to go to patients with actual speller
	interface}
\end{enumerate}
Minor comments:
\begin{enumerate}
	\item Measurements unit representation has been uniformized according to the
    SI standard throughout the text and in figures 5.4-5.6.
  \item Numerical citations have now consistently been replaced by text-style citations
    mentioning the author names where appropriate in the sentence.
  \item Contractions have been remove throughout the text.
\end{enumerate}

### 2024-11-19 chapters/wcble.tex
Quantify qualitative statements in wcble chapter

\Ac{eeg}-based visual oddball \acp{bci} establish communication with paralyzed
individuals by decoding \acp{erp} obtained from time-modulated flashing
stimuli (see~\cref{fig:bci/loop}).
These \acp{erp} consist of multiple \ac{erp}-components, peaks and troughs in
\ac{erp} waveform (see \cref{fig:bci/erp}).
These components each have their own distinct amplitude and latency.
\Ac{erp} component latency can reflect a wide range of cognitive,
neurophysiological, and clinical properties, that can be confounding factors in
the development of \acp{bci}~\cite{Luck2014}.
Latency is often assessed as the timing of an \ac{erp} component peak, obtained
after averaging over trials.
Latencies of non-averaged, single-trial \acp{erp} should capture even more
relevant information, yet they are very hard to measure due to the low \ac{snr} of
single-trial \acp{erp}.
Within-session variability of \ac{erp} latency, called \ac{erp} latency
\emph{jitter}, has been shown to influence \ac{bci}
accuracy~\cite{Thompson2012}.
This effect is especially prominent when bringing \ac{bci} development from the
lab to the clinical setting.
For instance, \textcite{Zisk2021} showed that \ac{als} results in significantly
increased \ac{erp} jitter\todo{qualitative}.
\textcite{Arico2014} showed that latency jitter negatively affects \ac{bci}
performance, and disproportionately so when the user is not directly gazing at
intended targets, such as is the case for individuals with \ac{sspgi}.
Therefore, properly accounting for \ac{erp} latency jitter can contribute to
the development of gaze-independent visual \acp{bci}.

@@

% ERROR BETWEEN HERE ==========================================================
100 target epochs containing the evoked potential and noise and 100
non-target epochs containing only noise were simulated.
The target epochs were jittered by setting the latency offset $l$
of each target epoch to a random latency drawn from a normal distribution with 0 mean and standard
deviations of respectively $\sigma=0.1,0.2,0.3$s.
The noise was scaled at 32 different noise levels, with the \ac{snr} of a
target epoch ranging from 0 to -31dB, with
\begin{equation}
  \text{SNR} = 10\log_{10}\frac{\var\left[s(t,0)\right]}{\var\left(\text{noise}\right)}
\end{equation}

@@

% AND HERE ====================================================================
As the \ac{snr} decreases, both \ac{cble} and \ac{wcble} can accurately
estimate evoked potential latencies up to a given threshold. After this
threshold, correlation between estimated and ground-truth latencies will start
decreasing.
However, we notice that this threshold occurs later for \ac{wcble} than for
\ac{cble} for all evaluated jitter levels, indicating that \ac{wcble} is more
robust in the presence of pink spatiotemporal noise.
At around -28dB, \ac{wcble} does not significantly outperform \ac{cble}
anymore, and both methods get overtaken by the strong presence of noise.

@@

\Cref{fig:wcble/results/accuracy} shows that the classification accuracy of both \ac{tlda} and \ac{cble}
decreases with
$\sigma$ and \ac{snr}, but \ac{cble} outperforms \ac{tlda}.
This difference is particularly large\todo{qualitative} when $\sigma$ is higher ($\sigma=0.2$
and $\sigma=0.3$) and \ac{snr} is not yet high.
\Ac{wcble} always scored better than \ac{cble} and \ac{tlda}, except for
$\sigma=0.1$ and $\text{\ac{snr}} < 27 \text{dB}$, at which point both \ac{cble}
and \ac{wcble} performed at chance leveltodo{qualitative}.
When considering the accuracy of \ac{wcble}, we see that it shows more of a
plateau as \ac{snr} increases\todo{qualitative}.

@@

The key finding of the simulation study presented
in \cref{sec:wcble/results} is that \ac{wcble} is robust to higher evoked
potential jitter and lower \ac{snr} than \ac{cble}\todo{qualitative}.
It could therefore be a better candidate to apply in \ac{erp} analysis and
decoding in those cases where jittered \ac{erp} data is of interest.
At a certain noise threshold\todo{qualitative}, both methods begin to fail, but \ac{wcble}
delivers higher accuracy at intermediate \ac{snr} levels than
\ac{cble}\todo{qualitative}, likely due to the iterative alignment process.

>>>>

\Ac{eeg}-based visual oddball \acp{bci} establish communication with paralyzed
individuals by decoding \acp{erp} obtained from time-modulated flashing
stimuli (see~\cref{fig:bci/loop}).
These \acp{erp} consist of multiple \ac{erp}-components, peaks and troughs in
\ac{erp} waveform (see \cref{fig:bci/erp}).
These components each have their own distinct amplitude and latency.
\Ac{erp} component latency can reflect a wide range of cognitive,
neurophysiological, and clinical properties, that can be confounding factors in
the development of \acp{bci}~\cite{Luck2014}.
Latency is often assessed as the timing of an \ac{erp} component peak, obtained
after averaging over trials.
Latencies of non-averaged, single-trial \acp{erp} should capture even more
relevant information, yet they are very hard to measure due to the low \ac{snr} of
single-trial \acp{erp}.
Within-session variability of \ac{erp} latency, called \ac{erp} latency
\emph{jitter}, has been shown to influence \ac{bci}
accuracy~\cite{Thompson2012}.
This effect is especially prominent when bringing \ac{bci} development from the
lab to the clinical setting.
For instance, \textcite{Zisk2021} showed that \ac{erp} jitter was significantly higher
in 6 \ac{als} patients (1350 ms$^2$) compared to neurotypical controls
(553 ms$^2$).
\textcite{Arico2014} showed that P3 latency jitter negatively correlated
($\rho=0.17$) with \ac{bci} accuracy in a covert GeoSpell interface.
Here the user does not gaze directly at the targets, such as is
the case for individuals with \ac{sspgi}.
Therefore, properly accounting for \ac{erp} latency jitter can contribute to
the development of gaze-independent visual \acp{bci}.

@@

% ERROR BETWEEN HERE ==========================================================
100 target epochs containing the evoked potential and noise and 100
non-target epochs containing only noise were simulated.
The target epochs were jittered by setting the latency offset $l$
of each target epoch to a random latency drawn from a normal distribution with 0 mean and standard
deviations of respectively $\sigma=0.1,0.2,0.3$s.
The noise was scaled at 32 different noise levels, with the \ac{snr} of a
target epoch ranging from 0 to -31 dB, with
\begin{equation}
  \text{SNR} = 10\log_{10}\frac{\var\left[s(t,0)\right]}{\var\left(\text{noise}\right)}
\end{equation}

@@

% AND HERE ====================================================================
As the \ac{snr} decreases, both \ac{cble} and \ac{wcble} can accurately
estimate evoked potential latencies up to a given threshold. After this
threshold, correlation between estimated and ground-truth latencies will start
decreasing.
However, we notice that this threshold occurs later for \ac{wcble} than for
\ac{cble} for all evaluated jitter levels, indicating that \ac{wcble} is more
robust in the presence of pink spatiotemporal noise.
At around -28 dB, \ac{wcble} does not significantly outperform \ac{cble}
anymore, and both methods get overtaken by the strong presence of noise.

@@

\Cref{fig:wcble/results/accuracy} shows that the classification accuracy of all
three methods decreases with $\sigma$ and \ac{snr}, yet \ac{wcble} outperforms
both \ac{cble} outperforms \ac{tlda} since its performance declines less steeply.
This difference is particularly large when $\sigma$ is higher
($\sigma=0.2$ and $\sigma=0.3$ s) and \ac{snr} is not yet high, with accuracies of respectively
77.50\% for \ac{wcble}, 17.00\% for \ac{cble} and 9.50\% for \ac{tlda} when
$\sigma=0.3$ and \ac{snr} = -23 dB.
\Ac{wcble} always scored better than \ac{cble} and \ac{tlda}, except for
$\sigma=0.1$ and $\text{\ac{snr}} < 27 \text{dB}$, at which point both \ac{cble}
and \ac{wcble} performed below the chance level of 50\%.

@@

The key finding of the simulation study presented
in \cref{sec:wcble/results} is that \ac{wcble} is robust to higher evoked
potential jitter and lower \ac{snr} than \ac{cble}.
It could therefore be a better candidate to apply in \ac{erp} analysis and
decoding in those cases where jittered \ac{erp} data is of interest.
At a certain noise threshold (-27 dB for $\sigma=0.1$ s, -26 dB for
$\sigma=0.2$ s and -23 dB for $\sigma=0.3$ s), both methods begin to fail, but \ac{wcble}
delivers higher accuracy ($>75\%$) at intermediate \ac{snr} levels than
\ac{cble}, likely due to the iterative alignment process.
